Modal Logics in the Coq Proof Assistant

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write conclusion

The successful formalization of Scott's version of G\"odel's proof of
God's existence indicates that the \emph{embedding} of modal logics
into higher-order logics via the \emph{possible world semantics} is a
viable approach for interactive theorem proving within modal logics.
The minimalistic implementation of the embedding (Section
\ref{sec:modal}) takes special care to hide the underlying possible
world machinery from the user.  An inspection of the proof scripts in
Section \ref{sec:proof} shows that this goal has been achieved. The
user does not have to explicitly bother about worlds and their mutual
reachability; the provided tactics for modalities do the job for him/her.
Moreover, for subgoals that do not involve modalities, the user has all the
usual interactive tactics at his/her disposal.
%The implementation of hints to allow automation tactics to fully cope with the embedding remains for future work. Nevertheless, at this stage a tactic such as \texttt{firstorder} is able to prove all lemmas and theorems in Section \ref{sec:modal} fully automatically.

HOL Light, Isabelle

todo: automation, firstorder

The theological implications of the verified correctness of this proof
certainly depend on a critical discussion of the underlying concepts,
definitions and axioms, which is beyond the scope of this paper. 
%What
%can be said though is that our work presents further evidence that
%theism is at least not (obviously) irrational. 
Clearly, the application of theorem proving technology to the domain of
theoretical philosophy can --- as already pictured by Leibniz ---
be very fruitful for both areas. In fact, new results of
philosophical interest have been obtained by our
work~\cite{arXiv}. Moreover a first, basic infrastructure for
interactive and automated reasoning in higher-order modal logics has
been created and will be further improved in the future.

\paragraph{Acknowledgements:} we thanks Cedric Auger and Laurent Th\'ery, for their answers to our questions about Ltac in the Coq-Club mailing-list.

>>>>

The successful formalization of Scott's version of  G\"odel's
ontological argument indicates that the \emph{embedding} of  modal
logics into higher-order logics via the \emph{possible world
semantics} is a viable approach for fully interactive theorem proving
within modal logics. Our lightweight implementation of the embedding
(described in sections \ref{sec:Embedding} and \ref{sec:Tactics})
takes special care to hide the underlying possible world machinery
from the user. An inspection of the proof scripts in Section
\ref{sec:proof} shows that this goal has been achieved. The user does
not have to explicitly bother about worlds and their mutual
reachability; the provided tactics for modalities do the job for
him/her. Moreover, for subgoals that do not involve modalities, the
user has all the usual interactive tactics at his/her disposal.

Although fully automated (as opposed to interactive) theorem proving
is beyond the scope of this paper, it is worth mentioning that all
lemmas and theorems in sections \ref{sec:Embedding},
\ref{sec:SimpleExamples} and \ref{sec:BeyondK} could have been proven
automatically using \Coq's \texttt{firstorder} tactic. The
implementation of hints to allow \Coq's automatic tactics to take full
advantage of the embedding and the modal axioms still remains for
future work.

The theological implications of the verified correctness of Gödel's
proof certainly depend on a critical discussion of the underlying
concepts, definitions and axioms, which is beyond the scope of this
paper.  Clearly, the application of theorem proving technology to the
domain of theoretical philosophy can --- as already pictured by
Leibniz --- be very fruitful for both areas. We hope that the
infrastructure  that we have implemented for interactive and automated
reasoning in higher-order modal logics will be useful outside
philosophy as well.

\paragraph{Acknowledgements:} we thank Cedric Auger and Laurent Théry, 
for their answers to our questions about Ltac in the Coq-Club mailing-list.
